### 2017-03-31 paper/oopsla17-submission/introduction.tex
some overview

Recent work on type error \emph{localization}
%
\citep[\eg][]{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr}
%
frames localization as an optimization problem, seeking a set of
expressions to blame that minimize some cost function, \eg the number of
constraints in the blame set, or the ratio of ``good'' paths through the
constraint graph to ``bad'' paths.
%
However, even these techniques fail to distinguish between the two
underlined expressions in \autoref{fig:sumList}, it would be equally
valid to blame \emph{either} of them alone.
%
The |[]| on line 3 could be changed to |0|, or the |+| on line 4 could
be changed to either |@| (list append) or |::|, all of which would give
type-correct programs.
%
Thus, these state-of-the-art techniques are forced to either blame
\emph{both} locations, or choose one \emph{arbitrarily}.

@@

>>>>

Recent work on type error \emph{localization}
%
\citep[\eg][]{Zhang2014-lv,Loncaric2016-uk,Pavlinovic2014-mr},
%
determining which expression(s) to blame for the error,
frames localization as an optimization problem, seeking a set of
expressions to blame that minimize some cost function, \eg the number of
constraints in the blame set, or the ratio of ``good'' paths through the
constraint graph to ``bad'' paths.
%
However, even these techniques fail to distinguish between the two
underlined expressions in \autoref{fig:sumList}, it would be equally
valid to blame \emph{either} of them alone.
%
The |[]| on line 3 could be changed to |0|, or the |+| on line 4 could
be changed to either |@| (list append) or |::|, all of which would give
type-correct programs.
%
Thus, these state-of-the-art techniques are forced to either blame
\emph{both} locations, or choose one \emph{arbitrarily}.

@@

Overall, our results show that data-driven diagnosis is a practical
technique for localizing type errors by learning from past mistakes.
%
% We believe, furthermore, that with the right set of features our
% approach could be applied to other domains where there are multiple,
% equally valid explanations for an error.
% \ES{this sounds really dumb... need better phrasing}

### 2017-03-31 paper/oopsla17-submission/overview.tex
some overview

\paragraph{\textbf{Challenge (1):} How to represent programs?}

\paragraph{\textbf{Solution:} Use a \textbf{set} of feature vectors}

\paragraph{\textbf{Challenge (2):} What is the ``correct'' fix?}

\paragraph{\textbf{Solution:} Use the students fixes}

\paragraph{\textbf{Challenge (3):} What should be presented to the user?}

\paragraph{\textbf{Solution:} Rank predictions and present most likely}

>>>>

\paragraph{\textbf{Challenge (1): How to represent programs?}}

\paragraph{\textbf{Solution: Use a set of feature vectors}}

\paragraph{\textbf{Challenge (2): What is the ``correct'' expression to blame?}}
As we saw in \autoref{fig:sumList}, type errors admit many different
fixes to different expressions, each resulting in a different program.
%
How is the type checker to judge which expressions are more likely to be
incorrect than others?
%
Prior work has enlisted expert users to manually judge ill-typed
programs and determine the ``correct'' fix, but this approach does not
scale well to a dataset large enough to perform machine learning.
%
Furthermore, while the expert user has intimate knowledge of the type
system, she cannot know what the student \emph{intended} to do.

\paragraph{\textbf{Solution: Use the students' subsequent fixes}}
%
Rather than relying on expert users, we allow our students to do the
work of assigning blame themselves.
%
Software development is an iterative process and our students will
eventually fix their own programs, perhaps after multiple ill-fated
attempts.
%
For each ill-typed program in our dataset, we find the first
\emph{subsequent} program that type checks and declare that to be the
fix.
%
From this pair of an ill-typed program and its fix, we can extract a
\emph{diff} of the abstract syntax trees.
%
For example, suppose our student fixed the |sumList| program in
\autoref{fig:sumList} by replacing |[]| with |0|, the diff would include
only the |[]| expression.
%
Thus we would determine that the |[]| expression (and \emph{not} the
|+|) is to blame for the error.

\paragraph{\textbf{Challenge (3): What should be presented to the user?}}
We now have the tools to train a classifier to predict, for each
expression in an ill-typed program, whether it should be changed.
%
But this is not yet particularly suitable as \emph{user feedback}.
%
A recent survey of developers by \citet{Kochhar2016-oc} found that
developers are unlikely to examine more than around five potentially
erroneous locations before falling back to manual debugging.
%
Thus, we should limit our predictions to a select few to be presented to
the user.

\paragraph{\textbf{Solution: Rank expressions by ``confidence''}}
Luckily, many machine learning classifiers produce not only a predicted
label, but also a metric that can be interpreted as the classifier's
\emph{confidence} in its prediction.
%
Thus, we \emph{rank} each expression by the classifier's confidence that
it should be changed, and present only the top-$k$ predictions to the
user (in practice $k=3$).
